The Etiology of Cybersecurity

### 2020-07-23 reports/paper_0/main.tex
minor change
minor change in equation,add explanations in intro

If we consider a natural (real-world) phenomenon, such as gravity, we can think
of a scientific theory as an \emph{abstract} mathematical formula (or a set of
formulas) that defines axiomatically a certain (natural) phenomenon in a
specific algebraic structure. While such a theory can be proved to be correct
within the mathematical setting it has been conceived in (e.g. with respect to
other axioms of the algebraic structure where it has been postulated, or with
respect to other background theories), no proof can be carried out to verify
that the theory properly abstracts the phenomenon it describes (i.e. that the
theory is a sound or complete approximation of the real phenomenon).
Similarly, the results of the application of a scientific theory (i.e. its
predictions) can be verified in the abstract mathematical structure but how can
we be sure that the theory correctly characterizes the phenomenon it describes? 
Popper's famous ``demarcation principle'' draws a line between scientific and
pseudo-scientific theories by proposing that \emph{any} scientific theory can
be considered correct as long as no empirical experiment falsify it. For
example, the theory of gravity, in its mathematical setting, can predict how
objects attract each other, and those predictions can be tested by empirical
experiment. If the theory correctly abstracts the phenomenon, no empirical
experiment should falsify the predictions of that theory.

@@

Instead of reasoning of what makes a system secure or insecure,
we reason on what causes insecurities.
Our hypothesis is that cybersecurity attacks are only caused by 
the presence of errors in the design or implementation of a system.
For example, a design or implementation error in a sanitization function
may allow an attacker to perform SQL-injections. If errors are necessary
to enable cybersecurity attacks, a theory that predicts all potential 
errors (or source of errors) can be effectively used to predict the
security of a system. With our approach, a list of weaknesses emerges
from the mathematical formulation of the system and predicts 4 main classes
of weaknesses. Those classes of weaknesses are used to calculate
all the insecurity configurations of all the components of a system,
obtaining a precise estimation of all potential cybersecurity-related risks
in any given system. Our hypothesis can be falsified by means of experiments,
testing if all the predicted weaknesses are present in the system under
test, or testing if other (not predicted by our hypothesis) weaknesses are
present.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=.9\columnwidth]{system-agent.pdf}
	\caption{xample of system and agent structure}
	\label{fig:system-agent}
\end{figure}

@@

Based on these hypothesis, we can formulate the concepts of security and
insecurity (in the $\abftheory$-framework) as mathematical equations.  Let us consider a CPS
$S$, represented as a graph $G=\langle V,E\rangle$ where $V$ represents the
set of functional blocks and ports of $S$, and $E\subseteq V\times E$ is the
set of pairs representing the channels and connections (data flows) between
functional blocks. We define $R\subseteq V\times F$, where $F$ is the set
of all the requirements of $S$, and extend $G$ as $G'=\langle E'\times V'\rangle$
with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi: E'\rightarrow RCC$ (where $RCC$ is
the set of relations in the RCC) be
the total function associating an RCC relation to each edge in $G'$, and
$\Pi$ be the set of all different permutations of RCC relations over $E'$
(i.e. $\Pi=\{\langle\pi(e_0)=EQ,\ldots,\pi(e_n)=EQ\rangle,\ldots,\langle \pi(e_0)=DR,\ldots,\pi(e_n)=DR\rangle\}$). If $\sigma:\Pi\rightarrow\{0,1\}$ is an evaluation function
such that $\sigma(p)=1$ (where $p\in\Pi$) iff the input configuration is satisfiable with respect to
the logical theory defining the algebraic structure (mereotopology) and constraints
of the calculus RCC (otherwise $\sigma$ returns $0$), we can define
\begin{displaymath}
	I=\sum_{p\in\{\Pi\setminus\pi_{EQ}\}} \sigma(p)
\end{displaymath}
where $\pi_{EQ}\in\Pi$ is the output of the function $\pi$ that associates only EQ relations
to any $e\in E'$, and $I\in\mathbb{N}$ represents all the insecurity configurations
of the CPS $S$ where, at least, one of the RCC relations isn't EQ.
In other words, we consider $\pi_{EQ}$ as the only secure configuration, and 
we characterize security as follows.

>>>>

If we consider a natural (real-world) phenomenon, such as gravity, we can think
of a scientific theory as an \emph{abstract} mathematical formula (or a set of
formulas) that defines axiomatically a certain (natural) phenomenon in a
specific algebraic structure. While such a theory can be proved to be correct
within the mathematical setting it has been conceived in (e.g. with respect to
other axioms of the algebraic structure where it has been postulated, or with
respect to other background theories), no proof can be carried out to verify
that the theory properly abstracts the phenomenon it describes (i.e. that the
theory is a sound or complete approximation of the real phenomenon).
Similarly, the results of the application of a scientific theory (i.e. its
predictions) can be verified in the abstract mathematical structure but how can
we be sure that the theory correctly characterizes the phenomenon it describes? 
Popper's famous ``demarcation principle'' draws a line between scientific and
pseudo-scientific theories by proposing that \emph{any} scientific theory can
be considered correct (or, better, acceptable) as long as no empirical experiment falsify it. For
example, the theory of gravity, in its mathematical setting, can predict how
objects attract each other, and those predictions can be tested by empirical
experiment. If the theory correctly abstracts the phenomenon, no empirical
experiment should falsify the predictions of that theory.

@@

\fix{mr}{In other words, if (as Herley mentions in \autocite{herley}) 
the problem is in the way we reason on cybersecurity, we should
see it in the way we reply to simple cybersecurity (foundamental) questions.
Let us consider the foundamental question ``what is cybersecurity?'',
as in \autocite{hintikka}, we can translate the wh-question into an equivalent
existentially quantified formula. 
But if the formula is the following second-order formula: $\exists P. P(s)\rightarrow Secure(s)$,
(where we ask if there exists a predicate $P$ such that, given any arbitrary 
system $s$, when $P$ holds for $s$ then $s$ can be considered secure),
when the answer is (for example) ``if the system is confidential, the
system is secure'' the formula becomes tautological. In fact, $P$ and $Secure$ can 
be considered as semantically equivalent (security is confidentiality and vice-versa).}
Instead of starting from reasoning on what makes a system secure or insecure,
we reason on what causes insecurities.
\fix{mr}{We focus on insecurities only caused by the exploitation of
cybersecurity attacks, and we assume that achieving security means preventing all those attacks 
from being exploitable or exploited.}
Our hypothesis is that cybersecurity attacks are only caused by 
the presence of errors in the design or implementation of a system.
For example, a design or implementation error in a sanitization function
may allow an attacker to perform SQL-injections. If errors are necessary
to enable cybersecurity attacks, a theory that predicts all potential 
errors (or source of errors) can be effectively used to predict the
security of a system.
\fix{mr}{Therefore, 
a possible answer to the initial question can be that cybersecurity is the 
absence of attacks and the absence of errors implies cybersecurity.
This, however, raises another question: ``which are all the errors
in (the engineering of) a system that make it possible for a cybersecurity attack
to happen?''. More rigorusly, which scientific theory predicts all
possible errors that causes cybersecurity attacks? 

With our approach, a list of weaknesses emerges
from the mathematical formulation of the system and predicts 4 main classes
of weaknesses. Those classes of weaknesses are used to calculate
all the insecurity configurations of all the components of a system,
obtaining a precise estimation of all potential cybersecurity-related risks
in any given system. Our hypothesis can be falsified by means of experiments,
testing if all the predicted weaknesses are present in the system under
test, or testing if other (not predicted by our hypothesis) weaknesses are
present.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=.9\columnwidth]{system-agent.pdf}
	\caption{Example of system and agent structure}
	\label{fig:system-agent}
\end{figure}

@@

Based on these hypothesis, we can formulate the concepts of security and
insecurity (in the $\abftheory$-framework) as mathematical equations.  Let us consider a CPS
$S$, represented as a graph $G=\langle V,E\rangle$ where $V$ represents the
set of functional blocks and ports of $S$, and $E\subseteq V\times E$ is the
set of pairs representing the channels and connections (data flows) between
functional blocks. We define $R\subseteq V\times F$, where $F$ is the set
of all the requirements of $S$, and extend $G$ as $G'=\langle E'\times V'\rangle$
with $E'= E\cup R$ and $V'=V\cup F$. Let $\pi: E'\rightarrow RCC$ (where $RCC$ is
the set of relations in the RCC) be
the total function associating an RCC relation to each edge in $G'$, and
$\Pi$ be the set of all different permutations of RCC relations over $E'$
(i.e. $\Pi=\{\langle\pi(e_0)=EQ,\ldots,\pi(e_n)=EQ\rangle,\ldots,\langle \pi(e_0)=DR,\ldots,\pi(e_n)=DR\rangle\}$). If $\sigma:\Pi\rightarrow\{0,1\}$ is an evaluation function
such that $\sigma(p)=1$ (where $p\in\Pi$) iff the input configuration is satisfiable with respect to
the logical theory defining the algebraic structure (mereotopology) and constraints
of the calculus RCC (otherwise $\sigma$ returns $0$), we can define
\begin{displaymath}
	I=\sum_{p\in\{\Pi\setminus\pi_{eq}\}} \sigma(p)
\end{displaymath}
where $\pi_{eq}\in\Pi$ is the output of the function $\pi$ that associates only EQ relations
to any $e\in E'$, and $I\in\mathbb{N}$ represents all the insecurity configurations
of the CPS $S$ where, at least, one of the RCC relations isn't EQ.
In other words, we consider $\pi_{eq}$ as the only secure configuration.
